Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2018-06-10 hpmoon-asoco.tex
I've rewritten the abstract, please @JJ let me know if it more clear now, and if all your questions have been properly addressed :)

@@

  \textcolor{blue}{Real world optimization problems usually have many dimensions and
  several objectives, and are usually tackled using Multi-Objective Evolutionary Algorithms. The high dimensionality requires exceptional
  computing power, and the fact that several dimensions have to be
  searched at the same time add additional difficulty.}
% One of the most extended methods to deal with high-dimensional
% multi-objective optimization problems are the distributed
% co-evolutionary algorithms.
% ----------------- This sentence is meaningless. It should be
% substituted by other stating that co-evolutionary algorithms can
% deal with it for some reason
% Besides, we're no longer saying that they are co-evolutionary in the
% title, but "island-based selective operator application". This
% should go to the abstract too - JJ

%In these methods %which ones? coevolutionary or selective operator
                   %application? - JJ
%  each processor deals only with a part of the decision
%space, that is, the individuals belonging to the same sub-population,
%\textcolor{blue}{which are called {\em islands} in this context}, explore the same subset of decision variables. 
\textcolor{blue}{An approach to cope with these issues is to allocate parts of the decision space to different
processors. Thus, each processor only deals with the individuals of a sub-population (called
island in this context) that explores the part of the decision space allocated to the processor.}

However, only
decomposable problems could strictly be implemented in parallel this
way. \textcolor{blue}{A new approach, which we call {\em selective operator application} is presented here to deal with un-decomposable
problems:} % so we are saying that high-dimensional multi-objective
          % problems are a challenge, but we are presenting a solution
          % for non-decomposable problems. What's what? - JJ
% Still not addressed - JJ
instead of sharing parts
of individuals, each processor applies the variation operators to a
specific subset of the whole individual. % this solves... - JJ
% not addressed - JJ
We have previously proved
that extending the length of these subsets, and therefore, sharing
\textcolor{blue}{some search space between islands, implies an
  improvement in different performance metrics}.
% In this paper, we intend to prove that adapting the size of the
% overlapping section, since it improves *, is able to obtain better
% results than a non-adaptive strategy. In order to prove that, we
% compare - JJ
In this paper, a new method for automatically
adapting the size of the overlapping, or shared, sections of the
individuals has been compared to other collaborative evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm \textcolor{blue}{in high-dimensional multi-objective problems}. 
Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.   % Say
                                % something how it's relevant, and
                                % what you have to do if you have so
                                % many decision variables - JJ
\textcolor{blue}{This opens up a new challenge that to be addressed in the future.}
% This
                                % solves a problem or challenge we
                                % hadn't stated. - JJ Pablo: metido, mira a ver si queda frase memorable

@@

\textcolor{blue}{Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}}, % No they don't, and
                                % if they do, add a reference - JJ FERGU: added references
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration. %depending on the
                                %language, this time might not be
                                %significant at all, and will depend
                                %on the length of the chromosome, not
                                %the part of the chromosome you will
                                %be applying it to. Besides, this is
                                %not the critical part of the
                                %algorithm, the critical part is
                                %searching the solution space. 
Different authors have % this is only one possibility. You should talk
                       % about simple island models, too - JJ
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a % first
                                % reference to co-evolution in this
                                % paper. Is it still a paper about
                                % co-evolution? - JJ
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

We have chosen those metrics not only because they have been used
extensively, especially in some of the papers presented in Section 2
% Still section 2? Please create a reference and use it - JJ
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Wang09parallel,Martens13asynchronous},
but also because they cover different quality criteria. Mathematical
formulation of these metrics can be found in
\cite{Dorronsoro13superlinear}.

@@

Obtained results also show that when increasing the number of islands, the
automatic overlapping method ($A$) significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in automatic overlapping
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. \textcolor{blue}{Since our method divides the individuals into blocks of divisible size, aggregation methods after migration, such as the ones used in \cite{Kimovski15Parallel} or \cite{Dorronsoro13superlinear} can be also tested. But as we are dealing with complete individuals in each island, the possibility of using a real number, such as 0.3 (30\% overlapping) instead an integer value opens future experimental setups.}
For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending \textcolor{blue}{on population size, number of islands, or other values, or using some statistical parameter setting method, such as Taguchi's \cite{Keratmatpour2018Taguchi}}. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on the results.
% 
% More discussion should be added. Why is adaptive better? Is the high
% performance really needed? - JJ
% STILL more discussion should be needed. For instance, the last
% paragraph of results begs an explanation of the influence of diversity on
% results - JJ Pablo: extending this section

>>>>

\newenvironment{bluecolor}{\par\color{blue}}{\par}

@@

%Real world optimization problems usually have many dimensions and
%  several objectives, thus requiring exceptional computing power to be addressed.

 % The high dimensionality requires exceptional computing power, and the fact that several dimensions have to be searched at the same time add additional difficulty.

%Usually, distributed co-evolutionary algorithms are used to cope with these issues, by allocating parts of the decision space to different processors. Thus, each processor only deals with the individuals of a sub-population (called island in this context) that explores a specific part of the decision space.

\begin{bluecolor}
One of the most extended methods to deal with high-dimensional multi-objective optimization problems are the distributed co-evolutionary algorithms. %Pablo: reducing previous two paragraphs into one
However, only
decomposable problems could strictly be implemented in parallel this
way.

A new approach, which we call {\em selective operator application}, is used here to deal with un-decomposable
problems: % so we are saying that high-dimensional multi-objective
          % problems are a challenge, but we are presenting a solution
          % for non-decomposable problems. What's what? - JJ
% Still not addressed - JJ PABLO: rewritten almost all the abstract.
instead of sharing parts
of individuals, each processor applies the variation operators to a
specific subset of the whole individual. This avoids extra steps to recompose the complete individuals from other islands before evaluation. 
% this solves... - JJ
% not addressed - JJ Pablo: done now
%We have previously proved that extending the length of these subsets, and therefore, sharing some search space between islands, implies an improvement in different performance metrics.

 In this paper, we intend to prove that automatically setting the size of the
 overlapping section, since it focus the search between borders of the different search spaces, is able to obtain better
 results than using the same size independently of the number of islands. In order to prove that, we
 have compared it to other collaborative evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm in high-dimensional multi-objective problems.

Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables: it is better to distribute them the higher the dimensionality, and to increase the overlapping size if individuals are distributed in a large number of islands.
  % Say
                                % something how it's relevant, and
                                % what you have to do if you have so
                                % many decision variables - JJ PABLO: Done
This opens up a new challenge to be addressed in the future, by experimenting on new ways to calculate the overlapping sizes depending of different factors and parameters.
\end{bluecolor}

@@

\textcolor{blue}{Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}}, 
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration. %depending on the
                                %language, this time might not be
                                %significant at all, and will depend
                                %on the length of the chromosome, not
                                %the part of the chromosome you will
                                %be applying it to. Besides, this is
                                %not the critical part of the
                                %algorithm, the critical part is
                                %searching the solution space. 
Different authors have % this is only one possibility. You should talk
                       % about simple island models, too - JJ
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a % first
                                % reference to co-evolution in this
                                % paper. Is it still a paper about
                                % co-evolution? - JJ
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

We have chosen those metrics not only because they have been used
extensively, especially in some of the papers presented in Section \ref{sec:soa}
% Still section 2? Please create a reference and use it - JJ Pablo: done
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Wang09parallel,Martens13asynchronous},
but also because they cover different quality criteria. Mathematical
formulation of these metrics can be found in
\cite{Dorronsoro13superlinear}.

@@

Obtained results also show that when increasing the number of islands, the
automatic overlapping method ($A$) significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in automatic overlapping
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. \begin{bluecolor}Since our method divides the individuals into blocks of divisible size, aggregation methods after migration, such as the ones used in \cite{Kimovski15Parallel} or \cite{Dorronsoro13superlinear} can be also tested. But as we are dealing with complete individuals in each island, the possibility of using a real number, such as 0.3 (30\% overlapping) instead an integer value opens future experimental setups.
For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending on population size, number of islands, or other values, or using some statistical parameter setting method, such as Taguchi's \cite{Keratmatpour2018Taguchi}. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on the results.
\end{bluecolor}
% 
% More discussion should be added. Why is adaptive better? Is the high
% performance really needed? - JJ
% STILL more discussion should be needed. For instance, the last
% paragraph of results begs an explanation of the influence of diversity on
% results - JJ Pablo: extending this section
